---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# fable.bsts
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

This package provides a tidy R interface to the bsts forecasting procedure using [fable](https://github.com/tidyverts/fable). This package makes use of the [bsts package](https://cran.r-project.org/package=bsts) for R.

_While the `BSTS()` function works in the `fable::model()` framework, not all functionality from the package has been implemented._

## Completed

- Stationary and non-stationary trends
- Regression and trigonometric seasonalities

## In Progress

- Holiday models
- Exogenous regressors
- Monthly-annual cycle models
- User-specified priors

## Use

```{r}
# library(tidyverse)
# 
# source("data-raw/lax_passengers.R")
# data <- lax_passengers %>%
#   mutate(date = yearmonth(date))
#   
# data %>%
#   fabletools::model(
#     bsts = BSTS(passengers ~ intercept() + ar() + seasonal("1 day"))
#   ) %>%
#   fabletools::forecast(h = 48) %>%
#   autoplot()
```

### Specials implemented:

- `intercept()` - static intercept models
- `ar()` - autoregressive models
- `level()` - local and shared level models
- `trend()` - local linear, semilocal linear and Student local linear models
- `seasonal()` - seasonal regression models
- `trig()` - trigonometric seasonality models
