---
title: "Fitting Bayesian structural time series with the bsts R package"
author: "David Holt"
date: "12/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bsts)     # load the bsts package
library(tidyverse)
library(tsibble)
library(tsbox)
library(fable)
library(fable.bsts)
```

This is a mock-up of how fable-style formulas could be used to express state specifications for bsts models. It includes code from Steven Scott's article "Fitting Bayesian structural time series with the bsts R package."

I'm using the following functions as formula "specials":
  `trend(type = c("level", "local", "semilocal", "AR"), ...)`
  `season(nseasons, duration = NULL, ...)`
  `xreg(data, ...)`
  `trig(period, frequencies, ...)`
  `holiday(holidays, ...)`
  
This allows the specification of the state through a formula such as
  `response ~ trend("level") + season(24) + xreg(.) + holiday(holidays)`


## Example 1: Nowcasting

```{r nowcasting}
# DATA
data(iclaims)     # bring the initial.claims data into scope

# MODEL 1
ss <- AddLocalLinearTrend(list(), initial.claims$iclaimsNSA)
ss <- AddSeasonal(ss, initial.claims$iclaimsNSA, nseasons = 52)
model1 <- bsts(initial.claims$iclaimsNSA,
               state.specification = ss,
               niter = 1000)

# FORECAST MODEL 1
pred1 <- predict(model1, horizon = 12)

# MODEL 2
model2 <- bsts(iclaimsNSA ~ .,
               state.specification = ss,
               niter = 1000,
               data = initial.claims)

# MODEL 3
model3 <- bsts(iclaimsNSA ~ .,
               state.specification = ss,
               niter = 1000,
               data = initial.claims,
               expected.model.size = 5)  # Passed to SpikeSlabPrior.

# DATA - fable
# code that would transform iclaims into a wide tsibble
iclaims <- ts_tsibble(initial.claims) %>% 
  spread(key = id, value = value)

# MODELS 1, 2, 3 - fable
mbl <- model(iclaims,
             bsts_1 = BSTS(iclaimsNSA ~ trend("local") + season(52), niter = 1000),
             bsts_2 = BSTS(iclaimsNSA ~ trend("local") + season(52) + xreg(.)),
             bsts_3 = BSTS(iclaimsNSA ~ trend("local") + season(52) + xreg(., expected_size = 5)))

# FORECAST MODEL 1 - fable
fbl <- mbl %>%
  select(-bsts_2, -bsts_3) %>%     # only because models 2 and 3 aren't modeled in the article
  forecast(mbl, h = 12)
```

## Example 2: Long term forecasting

We assume data is set up properly for each. The mbl (model table) could be passed automatically to the `forecast()` function in the fable version, if the only object needed at the end was the table of forecasts.

```{r longterm}
# MODEL 1 
ss1 <- AddLocalLinearTrend(list(), sp500)
model1 <- bsts(sp500, state.specification = ss1, niter = 1000)
pred1 <- predict(model1, horizon = 360)

# MODEL 2
ss2 <- AddSemilocalLinearTrend(list(), sp500)
model2 <- bsts(sp500, state.specification = ss2, niter = 1000)
pred2 <- predict(model2, horizon = 360)

# MODELS 1, 2 - fable
mbl <- model(sp500,
             bsts_1 = BSTS(iclaimsNSA ~ trend("local"), niter = 1000),
             bsts_2 = BSTS(iclaimsNSA ~ trend("semilocal"), niter = 1000))
fbl <- forecast(mbl, h = 360)
```

## Example 3: Recession modeling using non-Gaussian data

```{r nonGaussian}
# MODEL
ss <- AddLocalLevel(list(),
                    sigma.prior = SdPrior(sigma.guess = .1,
                                          sample.size = 1,
                                          upper.limit = 1),
                    initial.state.prior = NormalPrior(0, 5))
ts.model <- bsts(nber ~ ., ss, data = gdp, niter = 20000,
                 family = "logit", expected.model.size = 10)

# MODEL - fable
mbl <- model(gdp,
  bsts = BSTS(
    nber ~ trend("level",
                 sigma_prior = SdPrior(sigma_guess = .1, sample_size = 1, upper_limmit = 1),
                 initial_state_prior = NormalPrior(0, 5)) + 
      xreg(., expected_size = 10), 
    niter = 20000, family = "logit"))
```

## Conclusion

```{r conclusion}
# bsts
ss <- AddSeasonal(ss, y, nseasons = 24)
# fable 1
mbl <- model(data, bsts = BSTS(value ~ season(24)))
# fable 2
mbl <- model(data, bsts = BSTS(value ~ season("1 day")))

# bsts
ss <- AddSeasonal(ss, y, nseasons = 7, season.duration = 24)
# fable 1
mbl <- model(data, bsts = BSTS(value ~ season(7, duration = 24)))
# fable 2
mbl <- model(data, bsts = BSTS(value ~ season("1 week", duration = "1 day")))
```
